# The NEWS file

A `NEWS` file should be included to keep track of changes to the code from one version to the next.
It can be a top level file or in the `inst/` directory.
Only one `NEWS` file should exist in the repository.

The following are acceptable locations and formats:

```{r, echo=FALSE, results='asis'}
tab <- data.frame(
  location = c(
    "`./inst/NEWS.Rd`",
    "`./inst/NEWS`",
    "`./inst/NEWS.md`",
    "`./NEWS.md`",
    "`./NEWS`"
  ),
  format = c(
    "$\\LaTeX$",
    "formatted text see `?news`",
    "markdown",
    "markdown",
    "formatted text see `?news`"
  )
)
knitr::kable(tab)
```

Specifics on formatting can be found on the help page for `?news`.
[_Bioconductor_](https://bioconductor.org/) uses the `NEWS` file to create the semi-annual release announcement.
It must include list elements and **cannot** be a plain text file.

An example format:

```
Changes in version 0.99.0 (2018-05-15)
+ Submitted to Bioconductor

Changes in version 1.1.1 (2018-06-15)
+ Fixed bug. Begin indexing from 1 instead of 2
+ Made the following significant changes
  o added a subsetting method
  o added a new field to database
```

After you install your package, the following can be run to see if the `NEWS` is properly formatted:

```{r, eval=FALSE}
utils::news(package="<name of your package>")
```

The output should look similar to the following:

```
Changes in version 1.1.1 (2018-06-15):

    o   Fixed bug. Begin indexing from 1 instead of 2

    o   Made the following significant changes
	o added a subsetting method
	o added a new field to database

Changes in version 0.99.0 (2018-05-15):

    o   Submitted to Bioconductor
```

If you get something like the following there are formatting errors that need to be corrected:

```
Version: 0.99.0
Date: 2018-05-15
Text: Submitted to Bioconductor

Version: 1.1.1
Date: 2018-06-15
Text: Fixed bug. Begin indexing from 1 instead of 2

Version: 1.1.1
Date: 2018-06-15
Text: Made the following significant changes o added a subsetting
	method o added a new field to database
```
