# The NEWS file {#news}

A `NEWS` file informs the user of changes in the software that may impact user
expectations. It is highly recommended that the developer use plain and
concise language to communicate these changes.

A `NEWS` file should keep track of changes, in non-technical language, in the
code from one release version to the next. The file can reside in either
the package's top level directory or in the `inst/` directory. There should be
only one `NEWS` file per package.

The following are acceptable locations and formats:

```{r, echo=FALSE, results='asis'}
tab <- data.frame(
  location = c(
    "`./inst/NEWS.Rd`",
    "`./inst/NEWS`",
    "`./inst/NEWS.md`",
    "`./NEWS.md`",
    "`./NEWS`"
  ),
  format = c(
    "$\\LaTeX$",
    "formatted text see `?news`",
    "markdown",
    "markdown",
    "formatted text see `?news`"
  )
)
knitr::kable(tab)
```

Specifics on formatting can be found on the help page for `?news`.
[_Bioconductor_][] compiles `NEWS` files from all updated packages to create
the semi-annual release announcement.
`NEWS` files _must_ include list elements and **cannot** be plain text files.

Here is the version `1.24.0` `NEWS` file from `SummarizedExperiment`:

```
CHANGES IN VERSION 1.24.0
-------------------------

NEW FEATURES

    o Add 'checkDimnames' argument to SummarizedExperiment() constructor
      function

    o Add showAsCell() method for SummarizedExperiment objects.

SIGNIFICANT USER-VISIBLE CHANGES

    o Check the assay dimnames at SummarizedExperiment construction time:
      The SummarizedExperiment() constructor function now raises an error
      if one of the supplied assays has rownames and/or colnames that don't
      match those of the SummarizedExperiment object to construct.

```

**Note** that the version documented is the _forthcoming_ release version of
the package. For the developer, the `NEWS` file is a running log of all the
changes that lead up to the next release of the package.

The user can read the release notes of `SummarizedExperiment` version
`1.24.0` by running the following:

```{r, eval=FALSE}
utils::news(package = "SummarizedExperiment", Version == "1.24.0")
```

During development and after package re-installation, the `NEWS` file's
formatting can be checked with:

```{r, eval=FALSE}
utils::news(package="<name of your package>")
```

The function does not fail directly but it will show an output similar to the
one below if there are formatting errors that need to be corrected.
Note that there are bullet points embedded within the first `Text` field:

```
Version: 1.2.0
Date: 2018-07-15
Text: Made the following significant changes o added a subsetting
	method o added a new field to database

Version: 1.0.0
Date: 2018-06-15
Text: Fixed bug. Begin indexing from 1 instead of 2

Version: 0.99.0
Date: 2018-05-15
Text: Submitted to Bioconductor
```

