# Shiny apps {#shiny}

[Shiny apps][shiny-rstudio] can be submitted to _Bioconductor_ as software packages.

## Code organisation

Shiny apps are divided into two parts:

- User interface (UI)
- Server

Shiny apps submitted to _Bioconductor_ must store the <i class="fab fa-r-project"></i> code for both the UI and server code under the `R/` directory of the package.

The bulk of the package code should **not** be implemented directly within the `shinyApp()` function call.
Instead, internal package functions should be developped to produce and return individual components of the UI and server sides of the Shiny app.

## Running app

Functions in the package should return Shiny apps, but not **not** launch them.

In other words, the function `shiny::runApp()` should not be found anywhere in the package source code.
Instead, users should be left to call that function, using Shiny app objects returned by the package functions and options that control how the app is run (e.g., `launch.browser = TRUE`).

For instance, the package source code should look as follows:

```{r, eval=FALSE}
build_app <- function(...) {
  ui <- .build_ui(...)
  server <- .build_server(...)
  app <- shinyApp(ui = ui, server = server)
}
```

While the user's code should look as follows:

```{r, eval=FALSE}
app <- build_app(...)
shiny::runApp(app, ...)
```

## Building the package

_Coming soon: comments on build issues for packages that contain Shiny apps._

## Testing

The use of `# nocov start` and `# nocov end` is allowed to ignore part of the code that cannot be tested using traditional unit tests (e.g., observers).

For instance:

```{r, eval=FALSE}
# nocov start
observeEvent(input$example_input, {
    # <do something>
})
# nocov end
```

Use files `setup-*.R` in the subdirectory `tests/testthat/` to generate only once objects that are used repeatedly as input for the unit tests.

## Documentation

Man pages documenting functions that return Shiny apps should use the `interactive()` function to demonstrate usage of the app.

For instance, a typical 'Example' section for such a man page should look as follows:

```{r, eval=FALSE}
library(MyShinyPackage)

app <- build_app(...)

if (interactive()) {
  shiny::runApp(app, ...)
}
```

## Review

_Coming soon: comments on reviewing packages that contain Shiny apps (e.g., points raised during the review process)._
